Neural networks for solving regression and classification problems

### 2022-11-14 doc/paper.tex
heatmap text

From FIGURE MOMENTUM v CONSTANT we can observe that adding momentum reduces the amount of epochs necessary to reach a lower MSE score. Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate. Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs. We can also observe that the different schedulers that they all reach REASONABLE MSE within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data. Looking at TIME TABLE FOR ALL SCHEDULERS, BEST took the least time, beating out our baseline constant scheduler by TIME DIFF. Surprisingly, RMS prop performed more poorly than expected.

Before continuing, it is interesting to note that all the schedulers work best with a combination of similar eta and lambda values when gridsearched for achieving the minimal MSE. It is also interesting that they in the long term reach more or less the same error as can be observed in HEATMAP FIG. This could be because given more epochs, the other schedulers are given more time to catch up as the best performing schedulers, such as Adam, only decrease their MSE slightly. However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant schedulers.

However, when trained for 500 epochs, Adam with optimal parameters does not yield any improvement in MSE. Scikits Adam scheduler does perform better, further decreasing the MSE slightly. Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model. Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of ADAM FINAL MSE and Scikits final MSE of SCIKIT FINAL MSE both falling short of the analytical MSE of 0,003. In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions. This result might be due to 500 epochs not being enough training time.

>>>>

From FIGURE MOMENTUM v CONSTANT we can observe that adding momentum reduces the amount of epochs necessary to reach a lower MSE score. Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate.

- GD vs SGD

Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs. We can also observe that the different schedulers that they all reach REASONABLE MSE within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data. Looking at TIME TABLE FOR ALL SCHEDULERS, BEST took the least time, beating out our baseline constant scheduler by TIME DIFF. Surprisingly, RMS prop performed more poorly than expected.

An interesting insight from FIGURE NUM is that all the schedulers prefer high learning rates and regularization. As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time. With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization. For this specific plot, momentum does not seem to be an important parameter. Both the momentum and adagrad with momentum heatmaps are identical to their momentumless counterparts, constant and adagrad, implying that the best gridsearched momentum is 0. That momentum is irrelevant in this instance is not a generalizable result, as momentum has played a key role in previous plots and results.

Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same. As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adam, starts slowing down the rate at which the MSE is decreased. However, as can be seen in HEATMAP FIG, Adam offers a 11.11 \% \footnote{$\frac{constant MSE}{adam MSE} * 100 = \frac{0.018}{0.016} * 100 = 11.11\%$} decrease in MSE compared to the constant scheduler. It is clear from the results thus far that Adam is the best scheduler for approximating the Franke function when using linear regression with gradient descent.

Furthermore, when trained for 5000 epochs, Adam with optimal parameters keeps improving the MSE. Scikit learn's Adam scheduler does perform better, offering a greater decrease the MSE. Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model, in addition to other optimizations present in the scikit learn implementation. Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of an MSE of 0.012 on the test data and Scikits final MSE of 0.004 (granted this is on the train data, as scikit learn does not provide the loss curve for the test data).  both falling short of the analytical MSE of 0,003. In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions. This result might be due to 5000 epochs not being enough training time, as the Franke function is a quite complex function with added noise.
